The Three-HITs Theorem\footnote{This work was partially supported by COST Action CA15123.}

### 2017-03-19 Paper/Paper.tex
Usage of coequalizers to simplify the presentation.

\newcommand{\idmap}[0]{\Boperator{Id}}
\newcommand{\refl}[0]{\Boperator{refl}}
\newcommand{\hpushout}[0]{\Boperator{hpushout}}
\newcommand{\hocolim}[0]{\Boperator{hocolim}}
\newcommand{\seg}[0]{\Boperator{seg}}
\newcommand{\inl}[0]{\Boperator{inl}}
\newcommand{\inr}[0]{\Boperator{inr}}
\newcommand{\glue}[0]{\Boperator{glue}}
\newcommand{\inc}[0]{\Boperator{inc}}
\newcommand{\com}[0]{\Boperator{com}}
\newcommand{\Type}[0]{\operatorname{\textsc{Type}}}
\newcommand{\pt}[0]{*}

@@

The goal of this paper is to formally justify this idea by showing that the higher inductive types defined in \cite{basoldhigher} can be generated from the interval, homotopy pushout and colimit.
More concretely, we will prove the following theorem.
\begin{theorem}[Three-HITs Theorem]
\label{thm:three-hit}
In Martin-L\"of type theory extended with an interval object, homotopy pushouts and colimits, we can interpret each higher inductive type from \cite{basoldhigher}. 
This means that for each HIT we can define a type with the same introduction, elimination and computation rules.
\end{theorem}
Also, this generalizes the results in \cite{doorn2016constructing,kraus2014general,kraus2016constructions} where the result is showed for truncations.
With this result the metatheory of higher inductive types could be simplified significantly, because rather than a general class, one only needs to check metatheoretical properties for three HITs.

In Section \ref{sec:prelims} we shall recall some of the required material for this paper.
More concretely, we give the syntax of higher inductive types, and using that we define the interval, homotopy pushout, and colimit.
Next we define the approximating sequence of a higher inductive type in Section \ref{sec:approximator}, and in Section \ref{sec:rules} we show that the colimit of this sequence satisfies the rules of the given HIT.
For this we need some lemmata which are proved in Section \ref{sec:Lemmata}.

@@

With all these notions we can give the introduction, elimination and computation rules of higher inductive types.
The introduction rules for $H$ as given in Definition \ref{def:hit} are
%\begin{minipage}{0.6\textwidth}
\[
c : A \> H \rightarrow H
\]
%\end{minipage}
%\begin{minipage}{0.4\textwidth}
\[
p_i : \prod (x : B_i \> H), t_i = r_i.
\]
%\end{minipage}
We also have an elimination rule for which we use the lifting of constructor terms.
\begin{equation*}
	\AxiomC{$\vdash Y : H \rightarrow \Type$}
	\noLine
	\UnaryInfC{$\vdash f : \prod (x : A \> H), \bar{A} \> Y \> x \rightarrow Y \> (c \> x)$}
%	\noLine
	\AxiomC{$\vdash q_i : \prod (x : B_i \> H) (h_x : \bar{B}_i \> Y \> x), \hatt_i =_{p_j \, x} \hatr_j$}
	\BinaryInfC{$\vdash H\frec(f, q_1, \ldots, q_n) : \prod (x : H), Y \> x$}
	\DisplayProof
\end{equation*}
Let us abbreviate $H\frec(f, q_1, \ldots, q_n)$ by $H\frec$.
The type $H$ also has computation rules for each point $t : A \> H$
\begin{equation*}
H\frec \> (c \> t) \equiv f \> t \> (\bar{A} \> H\frec \> t),
\end{equation*}
and for each $a : B_i \> H$
\begin{equation*}
\apd \> H\frec \> (p_i \> a) \equiv q_i \> a \> (\bar{B}_i \> H\frec \> a).
\end{equation*}
Note that these equalities are definitional rather than propositional.

Let us now give some examples of higher inductive types which will be crucial in this paper.
The first one would be the interval
\lstset{language=Coq}
\begin{lstlisting}
Inductive $I^1$ :=
| $0$ : $I^1$
| $1$ : $I^1$
| $\seg$ : $0 = 1$
\end{lstlisting}
Note that for every type $A$ with inhabitants $x$ and $y$ we have a path $x = y$ iff we have a map $I^1$ to $A$ sending 0 and 1 to $x$ and $y$ respectively.
Let $\one$ be the unit type with point $\pt$.
Then we can define maps $\delta_0, \delta_1 : \one \rightarrow$ sending $\pt$ to 0 and 1 respectively.
Also, we define the boundary inclusion $\delta : \one + \one \rightarrow I^1$ by $\delta_0 + \delta_1$.

Next we define the homotopy pushout.
\lstset{language=Coq}
\begin{lstlisting}
Inductive $\hpushout$ $(A, B, C : \Type)$ $(f : A \rightarrow B)$ $(g : A \rightarrow C)$ :=
| $\inl$ : $B \rightarrow$ $\hpushout$ $A$ $B$ $C$ $f$ $g$
| $\inr$ : $C \rightarrow$ $\hpushout$ $A$ $B$ $C$ $f$ $g$
| $\glue$ : $\prod (a : A), \inl (f \> a) = \inr (g \> a)$
\end{lstlisting}
Note the similarities with the definition of the pushout \cite{mac2013categories}.
Lastly, we can define the homotopy colimit as a higher inductive type in much the same way.
\lstset{language=Coq}
\begin{lstlisting}
Inductive $\hocolim$ $(F : \mathbb{N} \rightarrow \Type)$ $(f : \prod (n : \mathbb{N}, F \> n \rightarrow F(n+1)))$ :=
| $\inc$ : $\prod (n : \mathbb{N}), F \> n \rightarrow$ $\hocolim$ $F$ $f$
| $\com$ : $\prod (n : \mathbb{N}) (x : F \> n), \inc \> n \> x = \inc \> (n+1) \> (f \> n \> x)$
\end{lstlisting}

@@

Before giving the definition, let us acquire inspiration by looking in more detail at how an inductive type is constructed.
Such a type $T$ is defined by a constructor $c : F \> T \rightarrow T$ with $F$ polynomial.
Equivalently, we can define the type with a signature which is a polynomial functor $F$, and then by Ad\'amek's theorem it is the following colimit.
\[
\xymatrix
{
	\zero \ar[r] & F \> \zero  \ar[r] & F(F \> \zero) \ar[r] & \ldots
}
\]
To understand what this does, let us assume that $F \> X = 1 + X$, so that $T = \mathbb{N}$.
This means we have two inclusions  $\xymatrix{1 \ar[r] & 1+X}$ and  $\xymatrix{X \ar[r] & 1+X}$, and we call them $0_C$ and $S_C$ respectively.
At every step we formally add for each $x : X$ a successor $S_C \> x$, and we add $0_C$.
Repeatedly applying this construction to the empty type $\zero$ gives the natural numbers $\mathbb{N}$.

One would like to construct higher inductive types in a similar fashion.
The first difference is that instead of starting with $\zero$, we start with all the nonrecursive constructors.
Also, in this construction, the functor $F$ is applied at every step.
Instead we will add terms built from the recursive constructors using previously constructed terms as arguments.
Since extra equalities might be present in the higher inductive type, we need to make identifications during the construction.
Rather than just adding points at every step, we also need to glue the specified paths.

To understand more precisely what should be done, let us consider an example.
\lstset{language=Coq}
\begin{lstlisting}
Inductive $\mathbb{N}_2$ :=
| $0$ : $\mathbb{N}_2$
| $S$ : $\mathbb{N}_2 \rightarrow \mathbb{N}_2$
| $p$ : $S(S \> 0) = 0$
\end{lstlisting}
The first approximation just has a constructor $0$, and after that we add a constructor for $S \> 0$ to obtain the second approximation.
In the third approximation, which we call $F' \> 3$, we found inhabitants $0$, $S \> 0$, and $S(S \> 0)$, and now we can make the first identification.
To do so, we take the following homotopy pushout
\[
\xymatrix
{
	\one + \one \ar[d]_{0 + S(S \> 0)} \ar[r]^-{\delta} & I^1 \ar[d]\\
	F' \> 3\ar[r] & F \> 3
}
\]
to obtain the actual third approximation $F \> 3$, and we continue our construction with that one.
Note that to glue during the $n$th step, we need refer to elements from the $(n-2)$th step.
So, in order to do the identification of the $n$th step, inhabitants from a previous approximation need to be used.

@@

Let us make this idea formal, and for that we start with a higher inductive type given as follows.
\lstset{language=Coq}
\begin{lstlisting}[label={lst:HIT}]
Inductive $H$ :=
| $c_{\nonrec}$ : $A_{\nonrec} \rightarrow H$
| $c_{\rec}$ : $A_{\rec} \> H \rightarrow H$
| $p_i$ : $\prod (x : B_i \> H), t_i = r_i$ $\quad (i = 1, \ldots, m)$
\end{lstlisting}
Note that the nonrecursive and recursive point constructors are separated in this definition.
The first approximation will be given using the nonrecursive constructors.
\lstset{language=Coq}
\begin{lstlisting}
Inductive $H_{\nonrec}$ :=
| $c_{\nonrec}'$ : $A_{\nonrec} \rightarrow H_{\nonrec}$
\end{lstlisting}

@@

\begin{lemma}
Suppose, we have a constructor term $t$ such that $x : F \> T \Vdash t : G \>T$ which uses at most $n$ constructors, and that we have a map $c_{\nonrec}' : A_{\nonrec} \rightarrow P$.
Then $t$ induces a map $\overline{t} : F \> P \rightarrow G \> H_{\Con}^n$ by replacing the constructors $c_{\nonrec}$ and $c_{\rec}$ by $c_{\nonrec}'$ and $c_{\rec}'$ respectively.
\end{lemma}

@@

Now we have sufficient to define the approximator.
During the construction one also needs to pay attention to coherency.
For example, in $P$ we have terms using $c_{\rec}$ at most $k$ times, so in $H_{\Approx} \> P$ we have duplicates of terms using $c_{\rec}$ less than $k$ times.
There will be several coherency problems during the proof, so the pushout will seem rather intimidating at first.

\begin{definition}
Let a higher inductive type $H$ be given as before, and let the type $H_{\Con}$ be defined as before.
Given are parameters $P$ and $Q$ with maps $i_l : H_{\Con} \> Q \rightarrow P$ and $i_r : I^1 \times B_i \> Q + Q \rightarrow P$.
Let us define the following maps.
\begin{align*}
\widetilde{i_l} &= i_l \comp \inn_1\\
g_1 &: ((\one + \one) \times B_i \> P) + H_{\Con} \> Q  + (I^1 \times B_i \> Q + Q)\rightarrow (I^1 \times B_i \> P) + P\\
g_1 &= (\inn_1 \comp \delta \times \idmap) + (\inn_2 \comp i_l) + (\inn_2 \comp i_r \circ \inn_1)\\
g_2 &: ((\one + \one) \times B_i \> P) + H_{\Con} \> Q + (I^1 \times B_i \> Q + Q) \rightarrow H_{\Con} \> P + (I^1 \times B_i \> P + P)\\
g_2 &= (\inn_1 \comp \overline{t_i} + \overline{r_i})+ (\inn_1 \comp H_{\Con} \> \widetilde{i_l}) + (\inn_2 \comp (I^1 \times B_i \> \widetilde{i_l}))
\end{align*}
Then we define the \emph{approximator} $H_{\Approx}$ of $H$ the following pushout.
\[
\xymatrix
{
	((\one + \one) \times B_i \> P) + H_{\Con} \> Q + (I^1 \times B_i \> Q) 
	\ar[d]_{g_2} 
	\ar[r]^-{g_1}
	& (I^1 \times B_i \> P) + P  
	\ar[d]^-{\inr}\\
	H_{\Con} \> P + (I^1 \times B_i \> P)  
	\ar[r]_-{\inl} 
	& H_{\Approx} \> P \> Q \> i_l \> i_r
}
\]
Note that this can be written as the homotopy pushout.
Often we will suppress the maps from the notation.
\end{definition}

The types $Q$ and $P$ represent $F(n-2)$ and $F(n-1)$ respectively, and the maps $i_l$ and $i_r$ are the inclusion maps into $F(n-1)$.
The new type is extended with constructors and paths given by $\inl \comp \inn_1$ and $\inr \comp \inn_1$ respectively.
Also, there are coherency conditions added and these are given by the second and third component of the left upper corner.
Their use will become clear in the Section \ref{sec:rules} when we construct the right introduction rules.
Now we can define the approximating sequence.

\begin{definition}
In the setting as described, we simultaneously define a sequence of approximations $F : \mathbb{N} \rightarrow \Type$ to $H$ and maps $f : \prod(n : \mathbb{N}), F \> n \rightarrow F(n+1)$ as follows
\begin{align*}
F \> 0 &= H_{\nonrec},\\
F \> 1 &= H_{\Approx} \> (F \> 0) \> \zero \> (\zero\frec (F \> 0)) \> (\zero\frec (F \> 0))\\
F(n+2) &= H_{\Approx} \> (F(n+1)) \> (F \> n) \> \inl \> \inr \qedhere
\end{align*}
For the maps maps $f \> n : F \> n \rightarrow F(n + 1)$, note that we always have the following sequence of maps
\[
\xymatrix@1
{
	P \ar[r] & (I^1 \times B_i \> P) + P \ar[r] & H_{\Approx} \> P \> Q \> i_l \> i_r.
}
\]
Taking $P$ to be $F \> n$, then we have $F (n+1) = H_{\Approx} (F \> n)$, and thus the composition gives the map $F \> n \rightarrow F(n + 1)$.
\end{definition}

@@

Now we will construct the map $A_{\rec} \> (\hocolim \> F \> f) \rightarrow \hocolim \> F \> f$, and by Lemma \ref{lem:colim_commute_poly} it suffices to make a map $\hocolim \> (A_{\rec} \comp F) \> (A_{\rec} \> f) \rightarrow \hocolim \> F \> f$.
For this we use the recursion rule of $\hocolim$, and we start with the following string of maps
\[
\xymatrix
{
		A_{\rec}(F \> n) \ar[r]^-{c_{\rec}'} 
		& H_{\rec}(F \> n) \ar[rr]^{\term \comp \iota_2} &
		& H_{\Con}(F \> n) \ar[r]^-{\inl}
		& H_{\Approx}(F \> n) = F(n + 1)
}
\]
Composing this map with $\inc$, gives maps $A_{\rec}(F \> n) \rightarrow \hocolim \> F \> f$ for all $n : \mathbb{N}$.

Next we need to show the commutativity of the following triangle.
\[
\xymatrix
{
	A_{\rec}(F \> n) \ar[rr]^{A_{\rec}(f \> n)} \ar[dr] & & A_{\rec}(F(n+1)) \ar[dl] \\
	& \hocolim \> F \> f &
}
\]
Before we continue, let us recall that inductive types are functors.
Suppose, we have an inductive type $T$ with a parameter $P$, then a function $f : P \rightarrow Q$ gives a function $T \> f : T \> P \rightarrow T \> Q$.
Let us start with the following rectangle
\[
\xymatrix
{
	A_{\rec}(F \> n) \ar[r]^-{c_{\rec}'} \ar[d]_{A_{\rec}(f \> n)} 
		& H_{\rec}(F \> n) \ar[rr]^{\term \comp \iota_2} \ar[d]_{H_{\rec}(f \> n)} &
		& H_{\Con}(F \> n) \ar[d]_{H_{\Con}(f \> n)}\\
	A_{\rec}(F(n+1)) \ar[r]_-{c_{\rec}'} 
		& H_{\rec}(F(n+1)) \ar[rr]_{\term \comp \iota_2} &
		& H_{\Con}(F(n+1))
}
\]
The left square commutes, because by definition of $H_{\rec} \> f$ we have
\[
H_{\rec} \>(f \> n) \> (c_{\rec}' x) = c_{\rec}' (A_{\rec} \> (f \> n) \> x).
\]
For a similar reason the right triangle commutes as well.
Hence, it suffices to show that the following diagram commutes.
\[
\xymatrix
{
	H_{\Con}(F \> n) \ar[d]_{H_{\Con} \> (f \> n)} \ar[r]^-{\inl}
		& F(n+1) \ar[d]^{f(n+1)}\\
	H_{\Con}(F(n+1)) \ar[r]_-{\inl}
		& H_{\Approx} \> (F(n+1)) \> (F \> n)
}
\]
Note that $H_{\Approx}(F(n+1)) = F(n+2)$ and $H_{\Approx}(F \> n) = F(n+1)$.
Hence, this diagram commutes, because of the pushout we used to define $H_{\Approx}$.

>>>>

\newcommand{\idmap}[0]{\Boperator{Id}}
\newcommand{\refl}[0]{\Boperator{refl}}

\newcommand{\coeq}[0]{\Boperator{coeq}}
\newcommand{\inC}[0]{\Boperator{inC}}
\newcommand{\glueC}[0]{\Boperator{glueC}}

\newcommand{\pcoeq}[0]{\Boperator{pcoeq}}
\newcommand{\inP}[0]{\Boperator{inP}}
\newcommand{\glueP}[0]{\Boperator{glueP}}

\newcommand{\hpushout}[0]{\Boperator{hpushout}}
\newcommand{\hocolim}[0]{\Boperator{hocolim}}
\newcommand{\seg}[0]{\Boperator{seg}}
\newcommand{\inl}[0]{\Boperator{inl}}
\newcommand{\inr}[0]{\Boperator{inr}}
\newcommand{\glue}[0]{\Boperator{glue}}
\newcommand{\inc}[0]{\Boperator{inc}}
\newcommand{\com}[0]{\Boperator{com}}
\newcommand{\Type}[0]{\operatorname{\textsc{Type}}}
\newcommand{\pt}[0]{*}

@@

The goal of this paper is to formally justify this idea by showing that the higher inductive types defined in \cite{basoldhigher} can be generated from three specific higher inductive types, namely coequalizers, path coequalizers, and homotopy colimit.
More concretely, we will prove the following theorem.
\begin{theorem}[Three-HITs Theorem]
\label{thm:three-hit}
In Martin-L\"of type theory extended with a coequalizers, path coequalizers and homotopy colimits, we can interpret each higher inductive type from \cite{basoldhigher}. 
This means that for each HIT we can define a type with the same introduction, elimination and computation rules.
\end{theorem}
Also, this generalizes the results in \cite{doorn2016constructing,kraus2014general,kraus2016constructions} where the result is showed for truncations.
With this result the metatheory of higher inductive types could be simplified significantly, because rather than a general class, one only needs to check metatheoretical properties for three HITs.

In Section \ref{sec:prelims} we shall recall some of the required material for this paper.
More concretely, we give the syntax of higher inductive types, and using that we define required types.
Next we define the approximating sequence of a higher inductive type in Section \ref{sec:approximator}, and in Section \ref{sec:rules} we show that the colimit of this sequence satisfies the rules of the given HIT.
For this we need some lemmata which are proved in Section \ref{sec:Lemmata}.
These are formalized in \textsc{Coq} \cite{CoqImpl} using the library from \cite{bauer2016hott}.

@@

With all these notions we can give the introduction, elimination and computation rules of higher inductive types.
The introduction rules for $H$ as given in Definition \ref{def:hit} are
\[
c : A \> H \rightarrow H,
\]
\[
p_i : \prod (x : B_i \> H), t_i = r_i.
\]
We also have an elimination rule for which we use the lifting of constructor terms.
\begin{equation*}
	\AxiomC{$\vdash Y : H \rightarrow \Type$}
	\noLine
	\UnaryInfC{$\vdash f : \prod (x : A \> H), \bar{A} \> Y \> x \rightarrow Y \> (c \> x)$}
%	\noLine
	\AxiomC{$\vdash q_i : \prod (x : B_i \> H) (h_x : \bar{B}_i \> Y \> x), \hatt_i =_{p_j \, x} \hatr_j$}
	\BinaryInfC{$\vdash H\frec(f, q_1, \ldots, q_n) : \prod (x : H), Y \> x$}
	\DisplayProof
\end{equation*}
Let us abbreviate $H\frec(f, q_1, \ldots, q_n)$ by $H\frec$.
The type $H$ also has computation rules for each point $t : A \> H$
\begin{equation*}
H\frec \> (c \> t) \equiv f \> t \> (\bar{A} \> H\frec \> t),
\end{equation*}
and for each $a : B_i \> H$
\begin{equation*}
\apd \> H\frec \> (p_i \> a) \equiv q_i \> a \> (\bar{B}_i \> H\frec \> a).
\end{equation*}
Note that these equalities are definitional rather than propositional.

Let us now give some examples of higher inductive types which will be crucial in this paper.
The first one identifies points in some type, and we call it the \emph{coequalizer}.
\lstset{language=Coq}
\begin{lstlisting}
Inductive $\coeq$ $(A, B : \Type)$ $(f, g : A \rightarrow B)$ :=
| $\inC$ : $B \rightarrow \coeq \> A \> B \> f \> g$
| $\glueC$ : $\prod (a : A), \inC \> (f \> a) = \inC \> (g \> a)$
\end{lstlisting}
The elimination rule for this type is as follows
\begin{equation*}
	\AxiomC{$\vdash Y : \coeq \> A \> B \> f \> g \rightarrow \Type$}
	\noLine
	\UnaryInfC{$\vdash i_Y : \prod (b : B), Y \> (\inC \> b)$}
	%	\noLine
	\AxiomC{$\vdash g_Y : \prod (a : A), \glueC_*(i_Y \> (f \> a)) = i_Y \> (g \> a)$}
	\BinaryInfC{$\vdash \coeq\frec(i_Y, g_Y) : \prod (x : \coeq \> A \> B \> f \> g), Y \> x$}
	\DisplayProof
\end{equation*}
Note the similarities with the definition of the coequalizer in category theory \cite{mac2013categories}.

Next we define a type which identifies two paths in some type which we shall call the \emph{path coequalizer}.
\lstset{language=Coq}
\begin{lstlisting}
Inductive $\pcoeq$ $(A, B : \Type)$ $(p : A \rightarrow \sum (b_1, b_2 : B), (b_1 = b_2) \times (b_1 = b_2))$ :=
| $\inP$ : $B \rightarrow \pcoeq \> A \> B \> f \> g$
| $\glueP$ : $\prod (a : A), \ap \> \inC \> (\pi_1(\pi_3(p \> a))) = \ap \> \inC \> (\pi_2(\pi_3(p \> a)))$
\end{lstlisting}
We will abbreviate $p_i \> a = \pi_i(\pi_3(p \> a))$ for $i = 1, 2$.
Note that this type has 2-paths, so it does not follow the given syntax.
Furthermore, the identified paths used in the parameter are obliged to have the same endpoints.
The introduction and computation rules are as expected, but we will give the elimination rule.
\begin{equation*}
	\AxiomC{$\vdash Y : \pcoeq \> A \> B \> f \> g \rightarrow \Type$}
	\noLine
	\UnaryInfC{$\vdash i_Y : \prod (b : B), Y \> (\inP \> b)$}
	%	\noLine
	\AxiomC{$\vdash g_Y : \prod (a : A), \apd \> i_Y \> (p_1 \> a) = \apd \> i_Y \> (p_2 \> a)$}
	\BinaryInfC{$\vdash \pcoeq\frec(i_Y, g_Y) : \prod (x : \pcoeq \> A \> B \> f \> g), Y \> x$}
	\DisplayProof
\end{equation*}

The last HIT we need is the homotopy colimit, and for that we again are inspired by the definition from category theory.
\lstset{language=Coq}
\begin{lstlisting}
Inductive $\hocolim$ $(F : \mathbb{N} \rightarrow \Type)$ $(f : \prod (n : \mathbb{N}, F \> n \rightarrow F(n+1)))$ :=
| $\inc$ : $\prod (n : \mathbb{N}), F \> n \rightarrow$ $\hocolim$ $F$ $f$
| $\com$ : $\prod (n : \mathbb{N}) (x : F \> n), \inc \> n \> x = \inc \> (n+1) \> (f \> n \> x)$
\end{lstlisting}
Now the elimination rule is as follows.
\begin{equation*}
	\AxiomC{$\vdash Y : \hocolim \> F \> f \rightarrow \Type$}
	\noLine
	\UnaryInfC{$\vdash i_Y : \prod (n : \mathbb{N}) (x : F \> n), Y \> (\inc \> n \> x)$}
	\noLine
	\UnaryInfC{$\vdash c_Y : \prod (n : \mathbb{N}) (x : F \> n), \com_*(i_Y \> n \> x) = i_Y \> (n+1) \> (f \> n \> x)$}
	\UnaryInfC{$\vdash \hocolim\frec(i_Y, c_Y) : \prod (x : \hocolim \> F \> f), Y \> x$}
	\DisplayProof
\end{equation*}

@@

Before giving the definition, let us acquire inspiration by looking in more detail at how an inductive type is constructed.
Such a type $T$ is defined by a constructor $c : A \> T \rightarrow T$ with $A$ polynomial.
Equivalently, we can define the type $T$ with a signature which is a polynomial functor $A$, and then by Ad\'amek's theorem $T$ is the following colimit.
\[
\xymatrix
{
	\zero \ar[r] & A \> \zero  \ar[r] & A(A \> \zero) \ar[r] & \ldots
}
\]
To understand what this does, let us assume that $A \> X = 1 + X$, so that $T = \mathbb{N}$.
This means we have two inclusions  $\xymatrix{1 \ar[r] & 1+X}$ and  $\xymatrix{X \ar[r] & 1+X}$, and we call them $0_C$ and $S_C$ respectively.
At every step we formally add for each $x : X$ a successor $S_C \> x$, and we add $0_C$.
Repeatedly applying this construction to the empty type $\zero$ gives the natural numbers $\mathbb{N}$.

One would like to construct higher inductive types in a similar fashion.
The first difference is that instead of starting with $\zero$, we start with all the nonrecursive constructors.
Also, in this construction, the functor $A$ is applied at every step.
Instead we will add terms built from the recursive constructors using previously constructed terms as arguments.
Since extra equalities might be present in the higher inductive type, we need to make identifications during the construction.
Rather than just adding points at every step, we also need to glue the specified paths.

To understand more precisely what should be done, let us consider an example.
\lstset{language=Coq}
\begin{lstlisting}
Inductive $\mathbb{N}_2$ :=
| $0$ : $\mathbb{N}_2$
| $S$ : $\mathbb{N}_2 \rightarrow \mathbb{N}_2$
| $p$ : $S(S \> 0) = 0$
\end{lstlisting}
We build a sequence $F$ of approximations.
The first approximation just has a constructor $0$, and after that we add a constructor for $S \> 0$ to obtain the second approximation.
Continuing this way, we ultimately arrive at the third approximation, which we call $F' \> 3$, where we found inhabitants $0$, $S \> 0$, and $S(S \> 0)$.
Now we can make the first identification, and to do so, we take the following homotopy pushout
\[
\xymatrix
{
	\one + \one \ar[d]_{0 + S(S \> 0)} \ar[r]^-{\delta} & I^1 \ar[d]\\
	F' \> 3\ar[r] & F \> 3
}
\]
to obtain the actual third approximation $F \> 3$, and we continue our construction with that one.
Note that to glue during the $n$th step, we need refer to elements from the $(n-2)$th step.
So, in order to do the identification of the $n$th step, inhabitants from a previous approximation need to be used.

@@

Let us make this idea formal, and for that we start with a higher inductive type given as follows.
\lstset{language=Coq}
\begin{lstlisting}[label={lst:HIT}]
Inductive $H$ :=
| $c$ : $A_{\nonrec} \rightarrow H$
| $c_{\rec}$ : $A_{\rec} \> H \rightarrow H$
| $p_i$ : $\prod (x : B_i \> H), t_i = r_i$ $\quad (i = 1, \ldots, m)$
\end{lstlisting}
Note that the nonrecursive and recursive point constructors are separated in this definition.
The first approximation will be given using the nonrecursive constructors.
\lstset{language=Coq}
\begin{lstlisting}
Inductive $H_{\nonrec}$ :=
| $c_{\nonrec}'$ : $A_{\nonrec} \rightarrow H_{\nonrec}$
\end{lstlisting}

@@

\begin{lemma}
Suppose, we have a constructor term $t$ such that $x : F \> T \Vdash t : G \>T$ which uses at most $n$ constructors, and that we have a map $c_{\nonrec}' : A_{\nonrec} \rightarrow P$.
Then $t$ induces a map $\overline{t} : F \> P \rightarrow G(H_{\Con}^n \> P)$ by replacing the constructors $c_{\nonrec}$ and $c_{\rec}$ by $c_{\nonrec}'$ and $c_{\rec}'$ respectively.
\end{lemma}

@@

Now we have sufficient to define the approximator.
Given is the following data
\begin{enumerate}
	\item Types $P, Q, R$;
	\item A map $j_Q : H_{\Con} \> Q \rightarrow P$;
	\item An inclusion $j_R : H_{\Con} \> R \rightarrow Q$;
	\item Paths $p_Q : \prod (x : Q), j_Q(\overline{t_i} \> x) = j_Q(\overline{r_i} \> x)$;
	\item Paths $p_R : \prod (x : R), j_R(\overline{t_i} \> x) = j_R(\overline{r_i} \> x)$.
\end{enumerate}
The type $H_{\Approx}$ depends on the given data, but for clarity we will suppress it from the notation.
We start with $H_{\Con} \> P$.

First, we need to guarantee that the added paths have the right endpoints.
Note that we have two maps $\overline{t_i}, \overline{r_i} : B_i \> P \rightarrow H_{\Con} \> P$, and that gives the first coequalizer.
\[
\xymatrix
{
	B_i \> P \ar@<-.5ex>[r]_-{\overline{r_i}} \ar@<.5ex>[r]^-{\overline{t_i}} & H_{\Con} \> P \ar[r]^-{\inC_1} & C_1
}
\]

However, this is not exactly the type we want due to coherency issues.
For example, in $P$ we have terms using $c_{\rec}$ at most $k$ times, so in $H_{\Approx} \> P$ we have duplicates of terms using $c_{\rec}$ less than $k$ times.

Remember that we we have $j_Q : H_{\Con} \> Q \rightarrow P$.
Now we define the type $C_2$ which is the coequalizer of the following two arrows.
\[
\xymatrix
{
	H_{\Con} \> Q \ar@<-.5ex>[rrr]_-{H_{\Con} \> (j_Q \circ \inn_1)} \ar@<.5ex>[rrr]^-{\inn_1 \circ j_Q} & & & H_{\Con} \> P \ar[rr]^-{\inC_1} & & C_1
}
\]
To explain what this does, let us assume for a moment that we have one construct $c$ with arity 1.
For $c' \> x$ with $x : Q$, the upper map is the inclusion of $H_{\Con} \> Q$ on the first coordinate of $H_{\Con} \> P$.
The lower map sends $c' \> x$ to $c' \> (j_Q \> (\inn_1 \> x))$, so it replaces $x$ by its image in $P$.

Another coherency we need is for the paths.
The paths use arguments from $B_i \> H$, and we want to prevent the presence of duplicates.
For this we define $C_3$ as the coequalizer of the following two arrows.
\[
\xymatrix
{
	B_i \> Q
	\ar@<-.5ex>[rrr]_-{H_{\Con} \> i \circ \overline{t_i}} \ar@<.5ex>[rrr]^-{\overline{t_i} \circ B_i(i \circ \inn_1)} 
	& & & 
	H_{\Con} \> P \ar[rrr]^-{\inC_2 \circ \inC_1}
	& & &
	C_2 
}
\]
Similarly, we define $C_4$ by replacing $\overline{t_i}$ by $\overline{r_i}$.

Lastly, we define $H_\Approx$, and for this we make one more identification.
Instead of identifying points, we will identify paths in this step.
Since also paths are added during the construction, there might also be duplicated paths.
The resulting type should be free, so these paths should be identified.
We define a map $q : B_i \> R \rightarrow \Sigma (x, y : P), (x = y) \times (x = y)$ sending $x : B_i \> R$ to the pair
\[
(\overline{t_i} \> x, \overline{r_i} \> x, \ap \> (i \circ \inn_1) \> (p_R \> x), p_Q \> (i_R \> x))
\]
Concluding the construction, we define $H_{\Approx}$ as $\pcoeq \> (B_i \> R) \> C_4 \> q$.

Now we can define the approximating sequence.

\begin{definition}
In the setting as described, we simultaneously define a sequence of approximations $F : \mathbb{N} \rightarrow \Type$ to $H$ and maps $f : \prod(n : \mathbb{N}), F \> n \rightarrow F(n+1)$ as follows
\begin{align*}
	F \> 0 &= H_{\nonrec},\\
	F \> 1 &= H_{\Approx} \> (F \> 0) \> \zero \> \zero \> \zero\frec \> \zero\frec \> \zero\frec \> \zero \frec \> \zero\frec\\
	F \> 2 &= H_{\Approx} \> (F \> 1) \> (F \> 0) \> \zero \> \inl \> \inr \> \zero\frec \> \zero\frec \> \zero\frec\\
	F(n+3) &= H_{\Approx} \> (F(n+2)) \> (F(n+1)) \> (F \> n) \> \inl \> \inr \> (f \> n) \> \overline{\glue} \> \overline{\glue} \qedhere
\end{align*}
where $\overline{\glue}$ is the map induced by the path $\glue$.
For the maps maps $f \> n : F \> n \rightarrow F(n + 1)$, note that we always have the following sequence of maps
\[
\xymatrix@1
{
	P \ar[r] & (I^1 \times B_i \> P) + P \ar[r] & H_{\Approx}.
}
\]
Taking $P$ to be $F \> n$, then we have $F (n+1) = H_{\Approx} (F \> n)$, and thus the composition gives the map $F \> n \rightarrow F(n + 1)$.
\end{definition}

@@

Now we will construct the map $A_{\rec} \> (\hocolim \> F \> f) \rightarrow \hocolim \> F \> f$, and by Lemma \ref{lem:colim_commute_poly} it suffices to make a map $\hocolim \> (A_{\rec} \comp F) \> (A_{\rec} \> f) \rightarrow \hocolim \> F \> f$.
For this we use the recursion rule of $\hocolim$, and we start with the following string of maps
\[
\xymatrix
{
		A_{\rec}(F \> n) \ar[r]^-{c_{\rec}'} 
		& H_{\rec}(F \> n) \ar[rr]^{\term \comp \inn_2} &
		& H_{\Con}(F \> n) \ar[r]^-{\inC}
		& H_{\Approx}(F \> n) = F(n + 1)
}
\]
Composing this map with $\inc$, gives maps $A_{\rec}(F \> n) \rightarrow \hocolim \> F \> f$ for all $n : \mathbb{N}$.

Next we need to show the commutativity of the following triangle.
\[
\xymatrix
{
	A_{\rec}(F \> n) \ar[rr]^{A_{\rec}(f \> n)} \ar[dr] & & A_{\rec}(F(n+1)) \ar[dl] \\
	& \hocolim \> F \> f &
}
\]
Before we continue, let us recall that inductive types are functors.
Suppose, we have an inductive type $T$ with a parameter $P$, then a function $f : P \rightarrow Q$ gives a function $T \> f : T \> P \rightarrow T \> Q$.
Let us start with the following rectangle
\[
\xymatrix
{
	A_{\rec}(F \> n) \ar[r]^-{c_{\rec}'} \ar[d]_{A_{\rec}(f \> n)} 
		& H_{\rec}(F \> n) \ar[rr]^{\term \comp \inn_2} \ar[d]_{H_{\rec}(f \> n)} &
		& H_{\Con}(F \> n) \ar[d]_{H_{\Con}(f \> n)}\\
	A_{\rec}(F(n+1)) \ar[r]_-{c_{\rec}'} 
		& H_{\rec}(F(n+1)) \ar[rr]_{\term \comp \inn_2} &
		& H_{\Con}(F(n+1))
}
\]
The left square commutes, because by definition of $H_{\rec} \> f$ we have
\[
H_{\rec} \>(f \> n) \> (c_{\rec}' x) = c_{\rec}' (A_{\rec} \> (f \> n) \> x).
\]
For a similar reason, the right square commutes as well.
Hence, it suffices to show that the following diagram commutes.
\[
\xymatrix
{
	H_{\Con}(F \> n) \ar[d]_{H_{\Con} \> (f \> n)} \ar[r]^-{\inC}
		& F(n+1) \ar[d]^{f(n+1)}\\
	H_{\Con}(F(n+1)) \ar[r]_-{\inC}
		& H_{\Approx} \> (F(n+1))
}
\]
Note that $H_{\Approx}(F(n+1)) = F(n+2)$ and $H_{\Approx}(F \> n) = F(n+1)$.
Hence, this diagram commutes, because of the coherency added in $C_2$.
